---
title: "Essay_Scoring"
output: pdf_document
---
Background





Question
Can short essay questions be automatically scored?


Data 
The training data set is available in the github repository in csv format. The score is in column 1 and the responses in column 2. The responses are parsed in the code.

     Level 6 - level 4 plus two additional responses
     Level 5 - level 4 plus one additional response
     Level 4 - include the following
     	spread out (open the space)
		cut/pass quickly (fake; use pivot; change speed, pathway, direction)
     Level 3 - one of the two responses (see level 4)
     Level 2 - incorrect response
     Level 1 - no response



Overview

The goal is to predict the score for each short essay response. This is the "first column variable (6,5,4,3,2,1) in the training set.


Code
```{r}
##Load libraries
library(caret)
library(tm)
library(ggplot2)
library(stats)
```

```{r}
##Read and clean data
answers <- read.csv("question.csv")
answers[,2] <- as.vector(answers[,2])
for(i in 1:299){
     answers[i,2] <- removePunctuation(answers[i,2])
     answers[i,2] <- tolower(answers[i,2])
}
```

```{r}
##Parse responses and explore data
review_answers <- answers[,2]
review_source <- VectorSource(review_answers)
corpus <- Corpus(review_source)
corpus <- tm_map(corpus, removeWords, stopwords("english"))
corpus <- tm_map(corpus, stemDocument)
ans <- DocumentTermMatrix(corpus)
ans2 <- as.matrix(ans)
ans3 <- cbind(answers[,1], ans2)
```

```{r}
##Create partition
set.seed (1229)
inTrain <- createDataPartition(ans3[, 1], p=0.60, list=F)
tempSet <- ans3[-inTrain, ] 
trainSet <- ans3[inTrain, ]
inTrain2 <- createDataPartition(tempSet[, 1], p=0.50, list=F)   
valSet <- tempSet[-inTrain2, ] 
testSet <- tempSet[inTrain2, ]   
```

```{r}
##Linear Model
ans_train <- data.frame(trainSet)
model1 <- lm(V1~., ans_train)
pred_train <- predict(model1,ans_train)
cor1 <- cor(round(pred_train), ans_train[,1])
cor1
```

Validation Set
```{r}
##Run linear model on ValSet
ans_val <- data.frame(valSet)
pred_val <- predict(model1,ans_val)
cor2 <- cor(round(pred_val), ans_val[,1])
cor2
```

Test Set
```{r}
##Run linear model on test_set
ans_test <- data.frame(testSet)
pred_test <- predict(model1,ans_test)
cor3 <- cor(round(pred_test), ans_test[,1])
cor3
```


Conclusion




Appendix A
Shows the relationship between prediction and actual scores for validation data

```{r}
plot(ans_val[,1],pred_val)

```

Appendix B
Shows the relationship between prediction and actual scores for test data
```{r}
plot(ans_test[,1],pred_test)

```

